\bf The usage of bootstrap method in shape-restricted regression

### 2023-08-13 Manuscript/manuscript.tex

The rest of the paper is organized as the follows. 
Section~\ref{Bootstrap Method for Constructing Confidence Intervals}
gives a review of bootstrap method for constructing confidence intervals
. Section~\ref{Shape Restricted Example} gives out a shape-restricted regression
example and the usage of bootstrap to build the pointwise confidence intervals
for certain time points. Section~\ref{Simulation Study} shows a simulation 
study to assess the performance of the methods. 
Section~\ref{Conclusion} concludes with a discussion.

@@

In some situations the regression is forced to be shape-restricted to improve 
the predictive performance and reduce overfitting if the underlying regression
function takes the specific form. The example is given from R package SemiPar.
And we fit the shape-restricted regression by the R package Splines2. Splines2
provides implementation of those spline basis functions via R language, and 
we use that to build a non-decreasing curve by i-spline that force the 
coefficients to be non-negative with degree of freedom = 6.

@@

We define a function \(y = 0.25(x - 0.9)\) and randomly generate 1000 points
\(x\) and its corresponding values \(y\) with added noise. The noise is added 
from a standard normal distribution with \(\mu = 0\) and standard deviation
\(1\). Let the paired values \((x, y\text{ with noise})\) be the dataset, 
and then we know the true values of \(y\) as well.

We fit shape-restricted i-splines regressions using different degrees of 
freedom (6, 10, 15) to the dataset. We select 10 equally spaced time points
from the dataset and perform bootstrap resampling 1000 times to estimate 
the \(95\%\) quantile pointwise confidence intervals for each of the selected
time points. We repeat this process 1000 times to see how many times these 
confidence intervals can contain the true result. The results are shown below:

\begin{table}[ht]
  \centering
  \caption{Result}
\begin{tabular}{|c|c|c|c|c|}
    \hline
\textbf{x} & \textbf{y} & \textbf{df\_6} & \textbf{df\_10} & \textbf{df\_15} \\
    \hline
    1 & 0.025 & 1 & 1 & 1 \\
    \hline
    4.22 & 0.831 & 1 & 1 & 1 \\
    \hline
    7.44 & 1.64 & 1 & 1 & 1 \\
    \hline
    10.7 & 2.44 & 1 & 1 & 1 \\
    \hline
    13.9 & 3.25 & 1 & 1 & 1 \\
    \hline
    17.1 & 4.05 & 0.999 & 1 & 1 \\
    \hline
    20.3 & 4.86 & 1 & 1 & 1 \\
    \hline
    23.6 & 5.66 & 1 & 1 & 1 \\
    \hline
    26.8 & 6.47 & 1 & 1 & 1 \\
    \hline
    30 & 7.28 & 1 & 1 & 1 \\
    \hline
  \end{tabular}
\end{table}

@@

The bootstrap method is a valid method to build pointwise confidence interval 
for shape-restricted regressions.

\bibliographystyle{asa}
\bibliography{citations}

>>>>

The rest of the paper is organized as the follows. Section~\ref{Bootstrap Method for Constructing Confidence Intervals} gives a review of bootstrap method for constructing confidence intervals. Section~\ref{Shape Restricted Example} gives out a shape-restricted regression example and the usage of bootstrap to build the pointwise confidence intervals for certain time points. Section~\ref{Simulation Study} shows a simulation study to assess the performance of the methods.
explores the case where a combination of the first two scenarios occurs. An  
adjusted bootstrap procedure is proposed as a working solution in this case.  
Section~\ref{Conclusion} concludes with a discussion.

@@

In some situations the regression is forced to be shape-restricted to improve the predictive performance and reduce overfitting if the underlying regression function takes the specific form. The example is given from R package SemiPar. And we fit the shape-restricted regression by the R package Splines2 based on the article \cite{wang2021shape}. Splines2 provides implementation of those spline basis functions via R language, and we use that to build a non-decreasing curve by i-spline that force the coefficients to be non-negative with degree of freedom = 6.

@@

We define three functions: \(y = 0.25(x - 0.9)\) , \(y = 0.02(x^3+1.2)\) , and \( y = 0.1 (5pnorm((x - 2) / 0.3) + 1)\) , and randomly generate 1000 points \(x\) and its corresponding values \(y\) with added noise. The noise is added from a standard normal distribution with \(\mu = 0\) and standard deviation \(1\). Let the paired values \((x, y\text{ with noise})\) be the dataset, and then we know the true values of \(y\) as well.

We fit shape-restricted i-splines regressions using degrees of freedom 6 to the dataset. We select 1, 5, 10, 15, 20, 25, 30 as the time points from the dataset and perform bootstrap resampling B = 1000 times to estimate the \(95\%\) quantile pointwise confidence intervals for each of the selected time points. We repeat this process 1000 times to see how many times these confidence intervals can contain the true result. The results are shown below:

\begin{table}[ht]
  \centering
  \caption{Results of function \(y = 0.25(x - 0.9)\) for different degrees of freedom}
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    \textbf{x} & \textbf{y} & \textbf{df\_6} & \textbf{df\_10} & \textbf{df\_15} \\
    \hline
    \multicolumn{5}{|c|}{B = 1000} \\
    \hline
    1 & 0.025 & 0.936 & 0.944 & 0.939 \\
    \hline
    5 & 1.025 & 0.960 & 0.958 & 0.948 \\
    \hline
    10 & 2.275 & 0.956 & 0.950 & 0.951 \\
    \hline
    15 & 3.525 & 0.959 & 0.954 & 0.954 \\
    \hline
    20 & 4.775 & 0.941 & 0.949 & 0.961 \\
    \hline
    25 & 6.025 & 0.953 & 0.946 & 0.942 \\
    \hline
    30 & 7.275 & 0.919 & 0.955 & 0.947 \\
    \hline
    \multicolumn{5}{|c|}{B = 2000} \\
    \hline
    1 & 0.025 & 0.936 & 0.944 & 0.939 \\
    \hline
    5 & 1.025 & 0.960 & 0.958 & 0.948 \\
    \hline
    10 & 2.275 & 0.956 & 0.950 & 0.951 \\
    \hline
    15 & 3.525 & 0.959 & 0.954 & 0.954 \\
    \hline
    20 & 4.775 & 0.941 & 0.949 & 0.961 \\
    \hline
    25 & 6.025 & 0.953 & 0.946 & 0.942 \\
    \hline
    30 & 7.275 & 0.919 & 0.955 & 0.947 \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Result of function  \(y = 0.02(x^3+1.2)\)}
\begin{tabular}{|c|c|c|c|c|}
    \hline
\textbf{x} & \textbf{y} & \textbf{df\_6} & \textbf{df\_10} & \textbf{df\_15} \\
    \hline
    1 & 0.025 & 0.893 & 0.940 & 0.959 \\
    \hline
    5 & 1.025 & 0.932 & 0.948 & 0.958 \\
    \hline
    10 & 2.275 & 0.937 & 0.948 & 0.941 \\
    \hline
    15 & 3.525 & 0.942 & 0.951 & 0.953 \\
    \hline
    20 & 4.775 & 0.939 & 0.951 & 0.957 \\
    \hline
    25 & 6.025 & 0.952 & 0.959 & 0.963 \\
    \hline
    30 & 7.275 & 0.947 & 0.943 & 0.943 \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Result of function \( y = 0.1 (5pnorm((x - 2) / 0.3) + 1)\) when B = 1000}
\begin{tabular}{|c|c|c|c|c|}
    \hline
\textbf{x} & \textbf{y} & \textbf{df\_6} & \textbf{df\_10} & \textbf{df\_15} \\
    \hline
    1 & 0.025 & 0.686 & 0.847 & 0.901 \\
    \hline
    5 & 1.025 & 0.279 & 0.800 & 0.852 \\
    \hline
    10 & 2.275 & 0.947 & 0.937 & 0.942 \\
    \hline
    15 & 3.525 & 0.937 & 0.953 & 0.961 \\
    \hline
    20 & 4.775 & 0.929 & 0.950 & 0.957 \\
    \hline
    25 & 6.025 & 0.953 & 0.924 & 0.922 \\
    \hline
    30 & 7.275 & 0.919 & 0.858 & 0.832 \\
    \hline
  \end{tabular}
\end{table}

@@

In all those functions and degrees of freedom we have given, the bootstrap method gives pointwise confidence intervals that are close to 95 \% . As the number of bootstrap samples \(B\) increases, the performance improves as expected, particularly for those points where the method initially showed lower performance. Therefore, we can conclude that the bootstrap method is a valid approach for constructing pointwise confidence intervals in the context of shape-restricted regressions.

\bibliographystyle{plain}
\bibliography{citations}
